MIRSVM: Multi-Instance Support Vector Machine with Bag Representatives

### 2017-07-22 ArticleRevision/article/article.tex
Updated Article based on Alberto's notes, still need confirmation about questions I asked him in the email.
> Page 1: complete
> Page 2: complete (minus VC dimension, but emailed Alberto for confirmation)
> Page 3: complete
> Page 8: complete
> Page 14: complete (minus the crossed out "imbalanced" word, emailed Alberto for confirmation)
> Page 15: complete
> Page 16: complete (with question about Wx p-value cut-off)
> Page 17: complete
> Page 18: complete (with edits to the text for formatting purposes)
> Page 19: complete
> Page 20: complete
> Page 21: complete

\begin{abstract}
\textit{Multiple-instance learning} (MIL) is a variation of \textit{supervised learning}, where samples are represented by labeled bags, each containing sets of instances. The individual labels of the samples within a bag are unknown, and labels are assigned based on a multi-instance assumption. One of the major complexities associated with this type of learning is the ambiguous relationship between a bag's label and the instances it contains. This paper proposes a novel support vector machine (SVM) multiple-instance formulation and presents an algorithm with a bag-representative selector that trains the SVM based on bag-level information, named MIRSVM. The contribution is able to identify instances that highly impact classification, i.e. bag-representatives, for both positive and negative bags, while finding the optimal class separation hyperplane. Unlike other multi-instance SVM methods, this approach eliminates possible class imbalance issues by allowing both positive and negative bags to have at most one representative, which constitute as the most contributing instances to the model. The experimental study evaluates and compares the performance of this proposal against $11$ state-of-the-art multi-instance methods over $15$ datasets, and the results are validated through non-parametric statistical analysis. The results indicate that bag-based learners outperform the instance-based and wrapper methods, as well as this proposal's overall superior performance against other multi-instance SVM models. 
\end{abstract}

@@

One of the major complexities associated with MIL is the ambiguity of the relationship between a bag label and the instances within the bag~\citep{Alpaydin2015}. This stems from the standard MI assumption, where the underlying distribution among instances within positive bags is unknown. There have been different attempts to overcome this complexity, such as ``flattenning" the MIL datasets, meaning instances contained in positive bags each adopt a positive label, allowing the use of classical supervised learning techniques~\citep{Ray2005}. This approach assumes that positive bags contain a significant number of positive samples, which may not be the case, causing the classifier to mislabel negative instances within the bag, decreasing the power of the MI model.

To overcome this, a different MIL approach was proposed, where subsets of instances are selected from positive bags for classifier training~\citep{Maron1998,Zhang2001}. One drawback of this type of method is that the resulting training datasets become imbalanced towards positive instances. Model performance further deteriorates when more instances are selected as subsets than needed~\citep{Carbonneau2016,Faria2017196,Vanwinckelen2016}. Our proposal aims to deal with this drawback by minimizing class imbalance. This is achieved by optimally selecting bag-representatives from both classes using support vector machines.

@@

To address the limitations presented by MIL algorithms, this paper proposes a novel support vector machine formulation with a bag-representative selector, called Multiple-Instance Representative Support Vector Machine (MIRSVM). SVMs are known to perform well when data is limited, therefore combining them with a bag-representative selector aims to remedy class imbalance caused by limited positive bag instances, without assuming the bags' internal distributions. The algorithm selects bag-representatives iteratively according to the standard MI assumption, ensuring known information, such as the negative bag labels, are fully utilized during the training process. The optimal separating hyperplane between bags is then found with respect to the bag-representatives using a Gaussian kernel and a quadratic programming solver. The optimal set of representatives is found when they have converged, or stopped changing from one iteration to the next. The algorithm does not assume any distribution of the instances and is not affected by the number of samples within a bag, making it applicable to a variety of contexts. The key contributions of this work include,
\begin{itemize}
\iitem Reformulating the traditional primal L$1$-SVM problem to optimize over bags, rather than instances, ensuring all the information contained within each bag is utilized during training, while defining bag representative selector criteria.

@@

This work is organized as follows. First, the notation used throughout the paper is provided, the MIL problem is formalized, and recent MIL methods, traditional support vector machines, and state-of-the-art multi-instance SVM methods are reviewed in Section~\ref{sec:background}. The primal L$1$-SVM is formulated with respect to optimizing over bags, the dual formulation is then derived along with conditions for optimality, and the proposed algorithm, MIRSVM, is listed and described in Section~\ref{sec:contribution}. Section~\ref{sec:experiments} presents the experimental environment, including results from various metrics, run-time, and non-parametric statistical analysis over $15$ benchmark MI datasets compared with $11$ state-of-the-art algorithms. Finally, Section~\ref{sec:conclusion} presents the conclusions of this contribution.

@@

\begin{table}
\caption{Summary of notation used throughout the paper}\label{tab:Notation}
\begin{tabularx}{\textwidth}{l@{\extracolsep{\fill}}l}
\hline\noalign{\smallskip}
Definition & Notation\\ 
\noalign{\smallskip}\hline\noalign{\smallskip}
Number of Bags & $n$ \\
Number of Instances & $m$ \\
Number of Input Attributes & $d$ \\
Set of Bags & $\mathcal{B} = \set{\mathcal{B}_1,\, \ldots,\, \mathcal{B}_n}$ \\
Bag Index Set & $I \in \ints_+^n$ \\
Input Space & $\bm{X} \in \mathbb{R}^{m \times d}$ \\
Bag Labels & $\bm{Y} \in \{-1,1\}^n$ \\
Input Instance $i$ from Bag $I$ & $\bm{x}_i = (x_{1}, \ldots, x_{d}),\, i \in I$ \\
Unknown Individual Instance Label $i$ & $y_i \in \set{-1,1}$ \\
Bag $I$ & $\mathcal{B}_I = \set{\bm x_i \st \forall i \in I}$ \\
\noalign{\smallskip}\hline\noalign{\smallskip}
Full Multi-Instance Training Dataset & $\mathcal{D} = \{(\mathcal{B}_1,Y_1), \ldots, (\mathcal{B}_n,Y_n)\}$ \\
\noalign{\smallskip}\hline
\end{tabularx}
\end{table}

@@

\citet{Zhou2009} proposed miGraph, which implicitely constructs a graph by deriving affinity matrices while using a graph kernel. It assumes the instances are not independently and identically distributed because of the nature of MI data. Bags imply relationships among the instances within it, and miGraph was designed to exploit them.

\section{Multiple-Instance Representative SVM}\label{sec:contribution}
MIRSVM is based on the idea of selecting representative instances from both positive and negative bags which are used to find an unbiased, optimal separating hyperplane. A representative is iteratively chosen from each bag, and a new hyperplane is formed according to the representatives until they converge. Based on the SMI hypothesis, only one instance in a bag is required to be positive for the bag to adopt a positive label. Due to the unknown distribution of instances within positive bags, MIRSVM is designed to give preference to negative bags during training, because their distribution is known, i.e. all instances are guaranteed to be negative. This is evident during the representative selection process, by taking the index of the maximum output value within each bag based on the current hyperplane using the following rule, $s_I = \text{argmax}_{i \in I} (\langle \bm w,\, \bm x_i \rangle + b),\, \forall I \in \set{1,\ldots,n}$. In other words, the most positive sample is chosen from each positive bag and the least negative sample is chosen from each negative bag (samples with the largest output value based on the current hyperplane), pushing the decision boundary  towards the positive bags. Equation~\eqref{eq:primalmirsvm} presents the primal MIRSVM optimization problem:
\begin{subequations} 
\label{eq:primalmirsvm}
\begin{alignat}{2}
\min\limits_{\bm w, b,\bm\xi}& {\,\,} \frac{1}{2}||\bm{w}||^2 + \frac{C}{n}\sum_I \xi_I, \tag{\ref{eq:primalmirsvm}}\\ 
\text{s.t.}& {\,\,} Y_I(\langle\bm{w},\bm x_{s_I}\rangle + b) \geq 1 - \xi_I,\, \forall I \in \set{1,\ldots,n}  \label{eq:primalmirsvm1}\\
{} & {\,\,} \xi_I \geq 0,\, \forall I \in \set{1,\ldots,n}, 
\end{alignat}
\end{subequations} 
where $\bm S_I$ is a set of the bag representatives' indices and $\bm x_{s_I}$ is the sample representative of bag $\mathcal{B}_I$. Note the variables in MIRSVMs formulation are the similar to those of the traditional SVM, except they are now representing each bag as an instance. Solving the optimization problem given in Equation~\eqref{eq:primalmirsvm} using a quadratic programming solver is a computationally expensive task due to the number of constraints, which scales by the number of bags $n$, as well as the calculation of the inner product between two $d$-dimensional vectors in constraint~\eqref{eq:primalmirsvm1}. The proposed solution for these problems was deriving and solving the dual of the optimization problem given by Equation~\eqref{eq:primalmirsvm}.

@@

Experiments were performed using $k$-fold cross validation, with $k = 10$, in order to objectively evaluate the models' performances and tune hyper-parameters. The data is separated fairly into $10$ equally sized sections where, at every iteration of the cross-validation loop, a section is held out as the test set, while the remainder of the data is used for training. These sets are stored and used by each algorithm, ensuring the cross-validation folds are controlled for each algorithm. This procedure ensures the model is not optimistically biased towards the full dataset and the algorithms are evaluated fairly over the same data in each fold. The tuning of the model during cross-validation includes finding the best penalty parameter, $C$, as well as the best shape parameter for the Gaussian radial basis function kernel, $\sigma$. The best hyper-parameters were chosen from the following $6 \times 6$ possible combination runs, shown in Equations~\eqref{eq:paramC} and~\eqref{eq:paramG}, referred to as~\eqref{eq:hyperparam}. 
\begin{subequations}
\label{eq:hyperparam}
\begin{align}
C \in  & \{0.1, 1, 10, 100, 1000, 10000\} \label{eq:paramC}\\
\sigma \in  & \{0.1, 0.5, 1, 2, 5, 10\} \label{eq:paramG}
\end{align}
\end{subequations}

@@

\subsection{Results \& Statistical Analysis}\label{subsec:results}
The classification performance was measured using five metrics: Accuracy~\eqref{eq:accuracy}, Precision \eqref{eq:precision}, Recall~\eqref{eq:recall}, Cohen's kappa rate~\eqref{eq:kappa}, and Area under ROC curve (AUC)~\eqref{eq:auc}. The Precision and Recall measures were reported because Accuracy alone can be misleading when classes are imbalanced, as is the case with the \textit{component} and \textit{function} datasets, which respectively have six and ten times as many negative instances than positive. Cohen's Kappa Rate and the AUC measures are used as complementary measures in order to evaluate the algorithms comprehensively~\citep{Bendavid2008}. Cohen's kappa rate, shown in Equation~\eqref{eq:kappa}, evaluates classifier merit according to the class distribution and ranges between -1 (full disagreement), 0 (random classification), and 1 (full agreement). The AUC metric highlights the trade-off between the true positive rate, or recall, and the false positive rate, as shown in Equation~\eqref{eq:auc}. The values of the true positive (TP), true negative (TN), false positive (FP), and false negative samples (FN) were first collected for each of the classifiers, then the metrics were computed using the equations shown in~\eqref{eq:metrics} on the $n^\prime$ bags of the test data. The run times (training and testing times) of each algorithm are also reported to analyze the usability and speed of each of the algorithms across differently sized datasets.

@@

The Holm and pairwise Wilcoxon rank-sum post-hoc~\citep{Gibbons2011,Holander1999} tests were then run for each of the metrics to compute multiple pairwise comparisons between the proposed algorithm and the state-of-the-art methods, investigating whether statistical differences exist among pairs of algorithms. Tables~\ref{tab:statacc},~\ref{tab:statprec},~\ref{tab:statrec},~\ref{tab:statkappa}, and~\ref{tab:statauc} show the $p$-values and Wilcoxon rank-sums for the Holm and Wilcoxon tests for $\alpha = 0.01$.
\newpage
\subsection{Accuracy}
\begin{table}[t!]
\caption{Accuracy}\label{tab:accResults}
\resizebox{0.95\textwidth}{!}{\begin{tabular}{l@{\extracolsep{\fill}}ccccccccccccc}
\noalign{\smallskip}\hline\noalign{\smallskip}
Datasets &MIRSVM & miGraph & MIBoost &MIOptimalBall &MIDD &MIWrapper &MISMO &MISVM &SimpleMI &TLC &Bagging &Stacking \\
\noalign{\smallskip}\hline\noalign{\smallskip}
suramin &0.8000 &\textbf{0.8462} &0.5000 &0.7250 &0.4250 &0.5000 &0.7250 &0.5000 &0.5000 &0.6000 &0.6650 &0.4615 &  \\
eastWest &\textbf{0.8000} &0.7000 &0.5000 &0.7250 &0.6125 &0.5000 &0.7125 &0.5625 &0.5000 &0.6000 &0.6000 &0.4500 &  \\
westEast &0.7500 &0.7500 &0.5000 &0.3750 &0.4500 &0.5000 &0.7375 &0.4125 &0.5000 &0.5625 &\textbf{0.9649} &0.6375 &  \\
musk1 &\textbf{0.9022} &0.8152 &0.5109 &0.7717 &0.8804 &0.5109 &0.7826 &0.7609 &0.5109 &0.8587 &0.8142 &0.8587 &  \\
musk2 &0.8218 &0.7426 &0.6139 &0.7723 &0.7228 &0.6139 &0.7030 &0.7129 &0.6139 &0.6238 &\textbf{0.8756} &0.6733 &  \\
webmining &0.8500 &0.8142 &0.8142 &0.7699 &0.8142 &0.8142 &0.8407 &0.6903 &0.8142 &0.8142 &\textbf{0.9358} &0.8053 &  \\
trx &0.8860 &0.8964 &0.8705 &\textbf{0.9016} &0.8808 &0.8705 &0.8705 &0.8705 &0.8705 &0.8756 &0.6450 &0.8860 &  \\
mutagenesis-atoms &0.7714 &0.7606 &0.6649 &0.6436 &0.7074 &0.6649 &0.6915 &0.6649 &0.6649 &\textbf{0.7766} &\textbf{0.7766} &0.7606 &  \\
mutagenesis-bonds &0.8252 &0.7872 &0.6649 &0.6915 &0.7713 &0.6649 &0.7979 &0.6649 &0.6649 &0.8351 &0.8351 &\textbf{0.8564} &  \\
mutagenesis-chains &\textbf{0.8411} &0.7926 &0.6649 &0.6702 &0.7766 &0.6649 &0.8351 &0.6649 &0.6649 &0.8404 &0.8404 &0.8351 &  \\
tiger &0.7750 &0.7950 &0.5000 &0.5000 &0.7100 &0.5000 &0.7200 &0.7550 &0.5000 &0.6650 &\textbf{0.8000} &0.7250 &  \\
elephant &\textbf{0.8300} &\textbf{0.8300} &0.5000 &0.5000 &0.7900 &0.5000 &0.8100 &0.8000 &0.5000 &0.8000 &0.5625 &0.8250 &  \\
fox &0.6550 &0.6300 &0.5000 &0.5000 &0.5800 &0.5000 &0.5250 &0.4750 &0.5000 &0.6450 &\textbf{0.8587} &0.6500 &  \\
component &\textbf{0.9366} &0.9153 &0.8649 &0.8696 &0.8780 &0.8649 &0.8968 &0.8703 &0.8649 &0.9358 &0.6000 &0.9355 &  \\
function &0.9523 &0.9405 &0.9155 &0.9138 &0.9193 &0.9155 &0.9376 &0.9195 &0.9155 &\textbf{0.9649} &0.6238 &0.9647 &  \\
\noalign{\smallskip}\hline\noalign{\smallskip}
Average &\textbf{0.8264} &0.8010 &0.6390 &0.6886 &0.7279 &0.6390 &0.7724 &0.6883 &0.6390 &0.7598 &0.7598 &0.7550 &  \\
Rank &\textbf{2.2000} &3.8667 &9.6000 &7.8667 &6.5667 &9.6000 &5.3333 &8.5667 &9.6000 &4.7000 &4.8667 &5.2333 &  \\
\noalign{\smallskip}\hline
\end{tabular}}
\centering
\resizebox{0.95\textwidth}{!}{\begin{tikzpicture}
\draw (1,0) -- (11,0);
\foreach \x in {1,2,3,4,5,6,7,8,9,10,11} {
\draw (\x, 0) -- ++(0,.1) node [above,scale=0.7] {\tiny \x};
\ifthenelse{\x < 11}{\draw (\x+.5, 0) -- ++(0,.03);}{}
}

@@

\foreach \x in {0,1,2,3,4,5,6,7,8,9,10,11} {
\draw (l\x) -| (c\x);};
\end{tikzpicture}}\vspace{-1.1em}
\captionof{figure}{Bonferroni-Dunn test for Recall}\label{fig:BonfDunnrec}\vspace{-0.5em}
\captionof{table}{Nemenyi, Holm, and Shaffer tests for Recall}\label{tab:statrec}
\scriptsize
\resizebox{0.95\textwidth}{!}{\begin{tabular}{lccccccccccc}
\noalign{\smallskip}\hline\noalign{\smallskip}
MIRSVM vs. &miGraph &MIBoost &MIOptimalBall &MIDD &MIWrapper &MISMO &MISVM &SimpleMI &TLC &Bagging &Stacking \\
\noalign{\smallskip}\hline\noalign{\smallskip}
Holm $p$-value &0.0125 &0.0063 &0.0167 &0.0050 &0.0071 &0.0100 &0.0045 &0.0500 &0.0056 &0.0250 &0.0083 \\\noalign{\smallskip}\hline\noalign{\smallskip}
Wilcoxon $p$-value &0.0060 & 0.2077 & 0.5614 & 0.4543 & 0.2077 & 0.6387 &0.1070 & 0.6603 & 0.5995 &0.1354 & 0.5721 \\
Wilcoxon R$^+$ &106.50 &83.000 &71.000 &74.000 &83.000 &69.000 &89.000 &60.000 &70.000 &87.000 &62.000 \\
Wilcoxon R$^-$ &13.500 &37.000 &49.000 &46.000 &37.000 &51.000 &31.000 &45.000 &50.000 &33.000 &43.000 \\
\noalign{\smallskip}\hline\noalign{\smallskip}
\end{tabular}}
\end{table}

>>>>

\begin{abstract}
\textit{Multiple-instance learning} (MIL) is a variation of \textit{supervised learning}, where samples are represented by labeled bags, each containing sets of instances. The individual labels of the instances within a bag are unknown, and labels are assigned based on a multi-instance assumption. One of the major complexities associated with this type of learning is the ambiguous relationship between a bag's label and the instances it contains. This paper proposes a novel support vector machine (SVM) multiple-instance formulation and presents an algorithm with a bag-representative selector that trains the SVM based on bag-level information, named MIRSVM. The contribution is able to identify instances that highly impact classification, i.e. bag-representatives, for both positive and negative bags, while finding the optimal class separation hyperplane. Unlike other multi-instance SVM methods, this approach eliminates possible class imbalance issues by allowing both positive and negative bags to have at most one representative, which constitute as the most contributing instances to the model. The experimental study evaluates and compares the performance of this proposal against $11$ state-of-the-art multi-instance methods over $15$ datasets, and the results are validated through non-parametric statistical analysis. The results indicate that bag-based learners outperform the instance-based and wrapper methods, as well as this proposal's overall superior performance against other multi-instance SVM models. 
\end{abstract}

@@

One of the major complexities associated with MIL is the ambiguity of the relationship between a bag label and the instances within the bag~\citep{Alpaydin2015}. This stems from the standard MI assumption, where the underlying distribution among instances within positive bags is unknown. There have been different attempts to overcome this complexity, such as ``flattening" the MIL datasets, meaning instances contained in positive bags each adopt a positive label, allowing the use of classical supervised learning techniques~\citep{Ray2005}. This approach assumes that positive bags contain a significant number of positive instances, which may not be the case, causing the classifier to mislabel negative instances within the bag, decreasing the power of the MI model.

To overcome this issue, a different MIL approach was proposed, where subsets of instances are selected from positive bags for classifier training~\citep{Maron1998,Zhang2001}. One drawback of this type of method is that the resulting training datasets become imbalanced towards positive instances. Model performance further deteriorates when more instances are selected as subsets than needed~\citep{Carbonneau2016,Faria2017196,Vanwinckelen2016}. Our proposal aims to deal with this drawback by minimizing class imbalance. This is achieved by optimally selecting bag-representatives from both classes using support vector machines.

@@

To address the limitations presented by MIL algorithms, this paper proposes a novel support vector machine formulation with a bag-representative selector, called Multiple-Instance Representative Support Vector Machine (MIRSVM). SVMs are known to perform well when data is limited, therefore combining them with a bag-representative selector aims to remedy class imbalance caused by limited positive bag instances, without assuming the bags' internal distributions. The algorithm selects bag-representatives iteratively according to the standard MI assumption, ensuring known information, such as the negative bag labels, are fully utilized during the training process. The optimal separating hyperplane between bags is then found with respect to the bag-representatives using a Gaussian kernel and a quadratic programming solver. The optimal set of representatives is found when they have converged, or stopped changing from one iteration to the next. The algorithm does not assume any distribution of the instances and is not affected by the number of instances within a bag, making it applicable to a variety of contexts. The key contributions of this work include,
\begin{itemize}
\iitem Reformulating the traditional primal L$1$-SVM problem to optimize over bags, rather than instances, ensuring all the information contained within each bag is utilized during training, while defining bag representative selector criteria.

@@

This work is organized as follows. Section~\ref{sec:background} presents the notation used throughout the paper, formalizes the MIL problem, and reviews recent MIL methods, traditional support vector machines, and state-of-the-art multi-instance SVM methods. Section~\ref{sec:contribution} presents the contribution: the primal L$1$-SVM formulation with respect to optimizing over bags, its derived dual formulation along with conditions for optimality, and the proposed algorithm, MIRSVM. Section~\ref{sec:experiments} presents the experimental environment, including results from various metrics, run-time, and non-parametric statistical analysis over $15$ benchmark MI datasets compared with $11$ state-of-the-art algorithms. Finally, Section~\ref{sec:conclusion} presents the conclusions of this contribution.

@@

\begin{table}[htbp]
\caption{Summary of notation used throughout the paper}\label{tab:Notation}
\begin{tabularx}{\textwidth}{l@{\extracolsep{\fill}}l}
\hline\noalign{\smallskip}
Definition & Notation\\ 
\noalign{\smallskip}\hline\noalign{\smallskip}
Number of Bags & $n$ \\
Number of Instances & $m$ \\
Number of Input Attributes & $d$ \\
Set of Bags & $\mathcal{B} = \set{\mathcal{B}_1,\, \ldots,\, \mathcal{B}_n}$ \\
Bag Index Set & $I \in \ints_+^n$ \\
Input Space & $\bm{X} \in \mathbb{R}^{m \times d}$ \\
Bag Labels & $\bm{Y} \in \{-1,1\}^n$ \\
Input Instance $i$ from Bag $I$ & $\bm{x}_i = (x_{1}, \ldots, x_{d}),\, i \in I$ \\
Unknown Individual Instance Label $i$ & $y_i \in \set{-1,1}$ \\
Bag $I$ & $\mathcal{B}_I = \set{\bm x_i \st \forall i \in I}$ \\
\noalign{\smallskip}\hline\noalign{\smallskip}
Full Multi-Instance Training Dataset & $\mathcal{D} = \{(\mathcal{B}_1,Y_1), \ldots, (\mathcal{B}_n,Y_n)\}$ \\
\noalign{\smallskip}\hline
\end{tabularx}
\end{table}

@@

\citet{Zhou2009} proposed miGraph, which implicitly constructs a graph by deriving affinity matrices while using a graph kernel. It assumes the instances are not independently and identically distributed because of the nature of MI data. Bags imply relationships among the instances within it, and miGraph was designed to exploit them.

\section{Multiple-Instance Representative SVM}\label{sec:contribution}
MIRSVM is based on the idea of selecting representative instances from both positive and negative bags which are used to find an unbiased, optimal separating hyperplane. A representative is iteratively chosen from each bag, and a new hyperplane is formed according to the representatives until they converge. Based on the SMI hypothesis, only one instance in a bag is required to be positive for the bag to adopt a positive label. Due to the unknown distribution of instances within positive bags, MIRSVM is designed to give preference to negative bags during training, because their distribution is known, i.e. all instances are guaranteed to be negative. This is evident during the representative selection process, by taking the index of the maximum output value within each bag based on the current hyperplane using the following rule, $s_I = \text{argmax}_{i \in I} (\langle \bm w,\, \bm x_i \rangle + b),\, \forall I \in \set{1,\ldots,n}$. In other words, the most positive instance is chosen from each positive bag and the least negative instance is chosen from each negative bag (instances with the largest output value based on the current hyperplane), pushing the decision boundary  towards the positive bags. Equation~\eqref{eq:primalmirsvm} presents the primal MIRSVM optimization problem:
\begin{subequations} 
\label{eq:primalmirsvm}
\begin{alignat}{2}
\min\limits_{\bm w, b,\bm\xi}& {\,\,} \frac{1}{2}||\bm{w}||^2 + \frac{C}{n}\sum_I \xi_I, \tag{\ref{eq:primalmirsvm}}\\ 
\text{s.t.}& {\,\,} Y_I(\langle\bm{w},\bm x_{s_I}\rangle + b) \geq 1 - \xi_I,\, \forall I \in \set{1,\ldots,n}  \label{eq:primalmirsvm1}\\
{} & {\,\,} \xi_I \geq 0,\, \forall I \in \set{1,\ldots,n}, 
\end{alignat}
\end{subequations} 
where $\bm S_I$ is a set of the bag representatives' indices and $\bm x_{s_I}$ is the instance representative of bag $\mathcal{B}_I$. Note the variables in MIRSVMs formulation are the similar to those of the traditional SVM, except they are now representing each bag as an instance. Solving the optimization problem given in Equation~\eqref{eq:primalmirsvm} using a quadratic programming solver is a computationally expensive task due to the number of constraints, which scales by the number of bags $n$, as well as the calculation of the inner product between two $d$-dimensional vectors in constraint~\eqref{eq:primalmirsvm1}. The proposed solution for these problems was deriving and solving the dual of the optimization problem given by Equation~\eqref{eq:primalmirsvm}.

@@

Experiments were performed using $10$-fold cross validation in order to objectively evaluate the models' performances and tune hyper-parameters. The data is separated fairly into $10$ equally sized sections where, at every iteration of the cross-validation loop, a section is held out as the test set, while the remainder of the data is used for training. These sets are stored and used by each algorithm, ensuring the cross-validation folds are controlled for each algorithm. This procedure ensures the model is not optimistically biased towards the full dataset and the algorithms are evaluated fairly over the same data in each fold. The tuning of the model during cross-validation includes finding the best penalty parameter, $C$, as well as the best shape parameter for the Gaussian radial basis function kernel, $\sigma$. The best hyper-parameters were chosen from the following $6 \times 6$ possible combination runs, shown in Equations~\eqref{eq:paramC} and~\eqref{eq:paramG}, referred to as~\eqref{eq:hyperparam}. 
\begin{subequations}
\label{eq:hyperparam}
\begin{align}
C \in  & \{0.1, 1, 10, 100, 1000, 10000\} \label{eq:paramC}\\
\sigma \in  & \{0.1, 0.5, 1, 2, 5, 10\} \label{eq:paramG}
\end{align}
\end{subequations}

@@

\subsection{Results \& Statistical Analysis}\label{subsec:results}
The classification performance was measured using five metrics: Accuracy~\eqref{eq:accuracy}, Precision \eqref{eq:precision}, Recall~\eqref{eq:recall}, Cohen's kappa rate~\eqref{eq:kappa}, and Area under ROC curve (AUC)~\eqref{eq:auc}. The Precision and Recall measures were reported because Accuracy alone can be misleading when classes are imbalanced, as is the case with the \textit{component} and \textit{function} datasets, which respectively have six and ten times as many negative bags than positive. Cohen's Kappa Rate and the AUC measures are used as complementary measures in order to evaluate the algorithms comprehensively~\citep{Bendavid2008}. Cohen's kappa rate, shown in Equation~\eqref{eq:kappa}, evaluates classifier merit according to the class distribution and ranges between -1 (full disagreement), 0 (random classification), and 1 (full agreement). The AUC metric highlights the trade-off between the true positive rate, or recall, and the false positive rate, as shown in Equation~\eqref{eq:auc}. The values of the true positive (TP), true negative (TN), false positive (FP), and false negative samples (FN) were first collected for each of the classifiers, then the metrics were computed using the equations shown in~\eqref{eq:metrics} on the $n^\prime$ bags of the test data, where $n^\prime = TP + FP + TN + FN$. The run times (training and testing times) of each algorithm are also reported to analyze the scalability and speed of each of the algorithms across differently sized datasets.

@@

The Holm (multiple) and Wilcoxon (pairwise) rank-sum post-hoc tests~\citep{Gibbons2011,Holander1999} were then run for each of the metrics to compute multiple and pairwise comparisons between the proposed algorithm and the state-of-the-art methods, investigating whether statistical differences exist among the algorithms' results. Tables~\ref{tab:statacc},~\ref{tab:statprec},~\ref{tab:statrec},~\ref{tab:statkappa}, and~\ref{tab:statauc} show the $p$-values for the Holm test for $\alpha = 0.05$, and the rank-sums and adjusted $p$-values for the Wilcoxon test.
\newpage
\subsection{Accuracy}
\begin{table}[t!]
\caption{Accuracy}\label{tab:accResults}
\resizebox{0.95\textwidth}{!}{\begin{tabular}{l@{\extracolsep{\fill}}ccccccccccccc}
\noalign{\smallskip}\hline\noalign{\smallskip}
Datasets &MIRSVM & miGraph & MIBoost &MIOptimalBall &MIDD &MIWrapper &MISMO &MISVM &SimpleMI &TLC &Bagging &Stacking \\
\noalign{\smallskip}\hline\noalign{\smallskip}
suramin &0.8000 &\textbf{0.8462} &0.5000 &0.7250 &0.4250 &0.5000 &0.7250 &0.5000 &0.5000 &0.6000 &0.6650 &0.4615 &  \\
eastWest &\textbf{0.8000} &0.7000 &0.5000 &0.7250 &0.6125 &0.5000 &0.7125 &0.5625 &0.5000 &0.6000 &0.6000 &0.4500 &  \\
westEast &0.7500 &0.7500 &0.5000 &0.3750 &0.4500 &0.5000 &0.7375 &0.4125 &0.5000 &0.5625 &\textbf{0.9649} &0.6375 &  \\
musk1 &\textbf{0.9022} &0.8152 &0.5109 &0.7717 &0.8804 &0.5109 &0.7826 &0.7609 &0.5109 &0.8587 &0.8142 &0.8587 &  \\
musk2 &0.8218 &0.7426 &0.6139 &0.7723 &0.7228 &0.6139 &0.7030 &0.7129 &0.6139 &0.6238 &\textbf{0.8756} &0.6733 &  \\
webmining &0.8500 &0.8142 &0.8142 &0.7699 &0.8142 &0.8142 &0.8407 &0.6903 &0.8142 &0.8142 &\textbf{0.9358} &0.8053 &  \\
trx &0.8860 &0.8964 &0.8705 &\textbf{0.9016} &0.8808 &0.8705 &0.8705 &0.8705 &0.8705 &0.8756 &0.6450 &0.8860 &  \\
mutagenesis-atoms &0.7714 &0.7606 &0.6649 &0.6436 &0.7074 &0.6649 &0.6915 &0.6649 &0.6649 &\textbf{0.7766} &\textbf{0.7766} &0.7606 &  \\
mutagenesis-bonds &0.8252 &0.7872 &0.6649 &0.6915 &0.7713 &0.6649 &0.7979 &0.6649 &0.6649 &0.8351 &0.8351 &\textbf{0.8564} &  \\
mutagenesis-chains &\textbf{0.8411} &0.7926 &0.6649 &0.6702 &0.7766 &0.6649 &0.8351 &0.6649 &0.6649 &0.8404 &0.8404 &0.8351 &  \\
tiger &0.7750 &0.7950 &0.5000 &0.5000 &0.7100 &0.5000 &0.7200 &0.7550 &0.5000 &0.6650 &\textbf{0.8000} &0.7250 &  \\
elephant &\textbf{0.8300} &\textbf{0.8300} &0.5000 &0.5000 &0.7900 &0.5000 &0.8100 &0.8000 &0.5000 &0.8000 &0.5625 &0.8250 &  \\
fox &0.6550 &0.6300 &0.5000 &0.5000 &0.5800 &0.5000 &0.5250 &0.4750 &0.5000 &0.6450 &\textbf{0.8587} &0.6500 &  \\
component &\textbf{0.9366} &0.9153 &0.8649 &0.8696 &0.8780 &0.8649 &0.8968 &0.8703 &0.8649 &0.9358 &0.6000 &0.9355 &  \\
function &0.9523 &0.9405 &0.9155 &0.9138 &0.9193 &0.9155 &0.9376 &0.9195 &0.9155 &\textbf{0.9649} &0.6238 &0.9647 &  \\
\noalign{\smallskip}\hline\noalign{\smallskip}
Average &\textbf{0.8264} &0.8010 &0.6390 &0.6886 &0.7279 &0.6390 &0.7724 &0.6883 &0.6390 &0.7598 &0.7598 &0.7550 &  \\
Rank &\textbf{2.2000} &3.8667 &9.6000 &7.8667 &6.5667 &9.6000 &5.3333 &8.5667 &9.6000 &4.7000 &4.8667 &5.2333 &  \\
\noalign{\smallskip}\hline
\end{tabular}}
\centering
\resizebox{0.95\textwidth}{!}{\begin{tikzpicture}
\draw (1,0) -- (11,0);
\foreach \x in {1,2,3,4,5,6,7,8,9,10,11} {
\draw (\x, 0) -- ++(0,.1) node [above,scale=0.7] {\tiny \x};
\ifthenelse{\x < 11}{\draw (\x+.5, 0) -- ++(0,.03);}{}
}

@@

\foreach \x in {0,1,2,3,4,5,6,7,8,9,10,11} {
\draw (l\x) -| (c\x);};
\end{tikzpicture}}\vspace{-1.1em}
\captionof{figure}{Bonferroni-Dunn test for Recall}\label{fig:BonfDunnrec}\vspace{-0.5em}
\captionof{table}{Holm and Wilcoxon tests for Recall}\label{tab:statrec}
\scriptsize
\resizebox{0.95\textwidth}{!}{\begin{tabular}{lccccccccccc}
\noalign{\smallskip}\hline\noalign{\smallskip}
MIRSVM vs. &miGraph &MIBoost &MIOptimalBall &MIDD &MIWrapper &MISMO &MISVM &SimpleMI &TLC &Bagging &Stacking \\
\noalign{\smallskip}\hline\noalign{\smallskip}
Holm $p$-value &0.0125 &0.0063 &0.0167 &0.0050 &0.0071 &0.0100 &0.0045 &0.0500 &0.0056 &0.0250 &0.0083 \\\noalign{\smallskip}\hline\noalign{\smallskip}
Wilcoxon $p$-value &0.0060 & 0.2077 & 0.5614 & 0.4543 & 0.2077 & 0.6387 &0.1070 & 0.6603 & 0.5995 &0.1354 & 0.5721 \\
Wilcoxon R$^+$ &106.50 &83.000 &71.000 &74.000 &83.000 &69.000 &89.000 &60.000 &70.000 &87.000 &62.000 \\
Wilcoxon R$^-$ &13.500 &37.000 &49.000 &46.000 &37.000 &51.000 &31.000 &45.000 &50.000 &33.000 &43.000 \\
\noalign{\smallskip}\hline\noalign{\smallskip}
\end{tabular}}
\end{table}

\subsection{Precision \& Recall}
Precision and recall are conflicting metrics that must be evaluated together in order to observe their behavior simultaneously, since they are both used to measure relevance. Tables~\ref{tab:preResults} and~\ref{tab:recResults} show the precision and recall results obtained by each algorithm. The precision and recall results for MIWrapper and SimpleMI indicate that they are unstable classifiers, exhibiting extreme variance in behavior, making them unsuitable for real-world applications. It is also interesting to analyze the performance on the mutagenesis datasets which have a larger number of positive bags than negative, where MISVM, MIBoost, MIWrapper, and SimpleMI predict all bags as negative. Additionally, while MISMO obtains unbiased results on these datasets, MIRSVM significantly outperforms it over both precision and recall, achieving a better trade-off.
